Towards understanding the central limit theorem by learning Python basics

### 2021-02-17 paper/article.tex
Add new section

@@

>>>>

% Own definitions (Zoltan)
\usepackage{listingsutf8,newtxtt,xcolor,graphicx,xurl}
\lstset{language=Python,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\color{magenta},
  stringstyle=\color{red},
  numberstyle=\color{blue}
  breaklines=true,
  showstringspaces=false,
  backgroundcolor=\color{yellow!30},
  ndkeywordstyle=\color{yellow},
  commentstyle=\color{green},
  identifierstyle=\color{black},
  literate=
  {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
  {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
  {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
  {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
  {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
  {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
  {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
  {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
  {Ã}{{\~A}}1 {ã}{{\~a}}1 {Õ}{{\~O}}1 {õ}{{\~o}}1
  {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
  {ű}{{\H{u}}}1 {Ű}{{\H{U}}}1 {ő}{{\H{o}}}1 {Ő}{{\H{O}}}1
  {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
  {€}{{\euro}}1 {£}{{\pounds}}1 {«}{{\guillemotleft}}1
  {»}{{\guillemotright}}1 {ñ}{{\~n}}1 {Ñ}{{\~N}}1 {¿}{{?`}}1,
  columns=fullflexible,
  keepspaces=true
}
\makeatletter
\def\lst@outputspace{{\ifx\lst@bkgcolor\empty\color{white}\else\lst@bkgcolor\fi\lst@visiblespace}}
\makeatother

@@

First author hold a series of supplementary lectures on probability theory two years ago, where second author
took part a student and illustrated several homework assignments by using Python programs.
Some of his illustrations were a bit complicated for beginners, so the idea came that
simplified versions of such programs could be helpful for many students.
But, instead of giving a problem in probability theory and asking for a solution, in this
project we reversed the logic: We revealed the Python code in advance, and wanted to know
what problem will be solved with its help.

%\begin{figure}
%\lstinputlisting[language=Python]{../progs/de_AT/1.py}
%\caption{The first program in its original form (in German)}
%\label{1.py}
%\end{figure}

The first example is shown in Fig.~\ref{main-py-window}. We kept the original version to highlight
the possibility to write special characters in the Python code (we used Python version 3),
including accented German letters. In the first set of problems each student had to try
the code on his or her own on the website \url{https://repl.it/languages/python3},
by paying attention on the correct indentation as well. We instructed all students with an
explanatory screenshot how the program should be started and the output obtained.

\begin{figure}
\begin{center}\includegraphics[scale=0.5]{../forms/de_AT/main-py-window}
\caption{An explanatory screenshot to help the students' first steps}
\label{main-py-window}
\end{center}
\end{figure}

In this introductory set the first question was to make the program work by running it.
To give a somewhat creative assignment we also asked how the output 3 can be changed to
the number 6. The outcome of this puzzle will be analyzed later in Section \ref{problem-sets}.

In fact, we created a couple of other programs in advance, but always waited for
all solutions turned in by the students before finalizing the next set of problems.
We \textit{did not} give any feedback on their solutions. Instead, if we felt that
some difficulties arose, then we tried to cover them in the next set of problems.
The reason behind this was to test if it is possible to create a set of problems
\textit{in advance} for a possible \textit{static} learning path. Even if our experiment
was not fully static, for a future project now we already have a static set of materials.

Here we emphasize that easy access to an online version of a Python interpreter
played a significant role in our project. We did not want to ask the students
to install anything on their computer. Instead, we tried to minimize the gap
between theory and practice. In fact, \url{repl.it} is not the only website
allowing the user to try his or her code freely---we were just
satisfied with its simple user interface. Later, to obtain graphical output in an easy way, we also
introduced how Sage worksheets at \url{cocalc.com} can be accessed---in this latter case,
however, accented letters did not look elegant.

Importantly, we simplified and minimized the code we used during the email course.
This pre-process required several discussions among the authors to avoid
the technical difficulties of the programming language, but still allow
obtaining simple outputs. We had to balance between readability, simplicity,
elegance and speed. We learned that Python itself could be improved to
have simpler constructions than the \texttt{for i in range(...)} loop for iteration.
At the end of day we were still satisfied with all of the code and the students'
solutions confirm that our choices were acceptable.

As an example of a successful balance, we mention that we heavily used Python's support
for Unicode variable names.
We denoted the sample space by $\Omega$ and a single outcome by $\omega$ since
these notations were already well-known for the students. On the other hand,
typing these Greek characters is not straightforward: one may need to copy-paste
them since they are not available on a German keyboard.

\section{The problem sets}
\label{problem-sets}
